AAnalyzing Student Use of Spacing and Interleaving Strategies in Interactions with GenAI-Powered Chatbots in Programming Courses

### 2025-01-15 CSEDU2025/Paper.tex
paper update

\cite{Bastani24} investigated the impact of GenAI, specifically GPT-4, on human
learning, with a focus on mathematics education in a high school. The problem
addressed was how the use of generative AI could affect the acquisition of new
skills, which is crucial for long-term productivity. The method involved a
controlled experiment with approximately one thousand students, who were exposed
to two GPT-4-based tutors: a simple tutor (GPT Base) and another with safeguards
designed to promote learning (GPT Tutor). The results showed that while access
to GPT-4 improved performance on practice exercises (48\% with GPT Base and
127\% with GPT Tutor), the removal of access to GPT Base led to a 17\% decrease
in student performance on exams. This suggests that unrestricted use of
\textit{GPT Base} could hinder learning. However, the GPT Tutor was able to
mitigate this negative effect.
\begin{table*}[htbp]
  \caption{Categories - adapted from \cite{Ghimire24}}
  \begin{center}
    \renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
    \begin{tabular}{p{3.8cm} p{6cm} p{4.5cm}} % Remove vertical bars
      \hline
      \textbf{Category} & \textbf{Description} & \textbf{Example of prompt} \\
      \hline
      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \textit{Would this code be ok? \{code\}} \\
      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. & \textit{What does it mean for a method to be static?} \\
      Student Correction (SC) & Prompts where the student corrects the chatbot. & \textit{The correct answer is B} \\
      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. & \textit{A class inherits from another write this code} \\
      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. & \textit{Give me the code for a selection sort} \\
      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). & \textit{A heap is a data structure appropriate for: \{options\}} \\
      Language Change (LC) & Prompts that request a change of idiom. & \textit{In Spanish} \\
      Uncategorized (U) & Prompts that do not fit into any of the above categories. & \textit{Thanks} \\
      \hline
    \end{tabular}
    \label{tab:categories}
  \end{center}
\end{table*}

@@

Figure \ref{fig:graph1} also indicates that approximately 5.8\% of the prompts
written by the students were classified as Uncategorized. These messages include
expressions of gratitude toward the chatbot, such as a simple '\textit{thanks}',
contextual statements like '\textit{I'm reviewing object orientation}', as well
as requests for additional exercises and summaries. Depending on the type of
investigation, there is opportunity for these messages to be further subdivided
into new categories.

Finally, 1.3\% of the messages were classified as Language Change. These prompts
were typically requests for the chatbot to switch languages, as seen in
conversation 9 in Figure \ref{fig:graph2}.

@@

On the other hand, it was observed that most students' messages concentrated on
categories such as Debugging Help and Conceptual Questions, which might suggest
that students were committed to understanding the concepts. However, the results
also indicated that students' use of study strategies such as Spacing and
Interleaving remains limited. Therefore, encouraging students to apply these
strategies intentionally could enhance their learning using GenAI.

>>>>

\cite{Bastani24} investigated the impact of GenAI, specifically GPT-4, on human
learning, with a focus on mathematics education in a high school. The problem
addressed was how the use of generative AI could affect the acquisition of new
skills, which is crucial for long-term productivity. The method involved a
controlled experiment with approximately one thousand students, who were exposed
to two GPT-4-based tutors: a simple tutor (GPT Base) and another with safeguards
designed to promote learning (GPT Tutor). The results showed that while access
to GPT-4 improved performance on practice exercises (48\% with GPT Base and
127\% with GPT Tutor), the removal of access to GPT Base led to a 17\% decrease
in student performance on exams. This suggests that unrestricted use of
\textit{GPT Base} could hinder learning. However, the GPT Tutor was able to
mitigate this negative effect.
\begin{table*}[htbp]
  \caption{Categories - adapted from \cite{Ghimire24}}
  \begin{center}
    \renewcommand{\arraystretch}{1.1} % Increase the spacing between rows
    \begin{tabular}{p{3.8cm} p{6cm} p{4.5cm}} % Remove vertical bars
      \hline
      \textbf{Category} & \textbf{Description} & \textbf{Example of prompt} \\
      \hline
      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \textit{Would this code be ok? \{code\}} \\
      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. & \textit{What does it mean for a method to be static?} \\
      Student Correction (SC) & Prompts where the student corrects the chatbot. & \textit{The correct answer is B} \\
      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. & \textit{A class inherits from another write this code} \\
      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. & \textit{Give me the code for a selection sort} \\
      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). & \textit{A heap is a data structure appropriate for: \{options\}} \\
      Language Change (LC) & Prompts that request a change of idiom. & \textit{In Spanish} \\
      Uncategorized (U) & Prompts that do not fit into any of the above categories. & \textit{Thanks} \\
      \hline
    \end{tabular}
    \label{tab:categories}
  \end{center}
\end{table*}

@@

Figure \ref{fig:graph1} also indicates that approximately 5.8\% of the prompts
written by the students were classified as Uncategorized. These messages include
expressions of gratitude toward the chatbot, such as a simple '\textit{thanks}',
contextual statements like '\textit{I'm reviewing object orientation}', as well
as requests for additional exercises and summaries. These Uncategorized messages
can be subdivided into new categories depending on the investigation. For
instance, requests for new exercises or summaries could be interpreted as a
metacognitive monitoring strategy.

Finally, 1.3\% of the messages were classified as Language Change. These prompts
were typically requests for the chatbot to switch languages (English to Spanish),
as seen in Figure \ref{fig:graph2}, conversation 9, message 2.

@@

The results also indicated that students' use of study strategies such
as Spacing and Interleaving remains limited. Therefore, encouraging students to
apply these strategies intentionally could enhance their learning using GenAI.
